Insight Into Lumbar Back Pain \subtitle{What the Lumbar Spine Tells About Your Life}

### 2014-08-27 draft.tex
added revised Results figure
begun related work section

Epidemiologists want to characterize the healthy aging process of the spine.
%%
To achieve this, they have to analyze the lumbar spine shape as well as the mentioned lifestyle factors.
%% -------------------------------
\section{\uppercase{Related Work}}
\label{sec:RelatedWork}
\noindent \com{Our own work (VIS, VMV, BVM). Sylvias paper with reference to the methods. \textbf{More information necessary here!}}

@@

  \includegraphics[width=0.8\textwidth]{figures/results}

>>>>

Epidemiologists want to characterize the healthy aging process of the spine.
%%
To achieve this, they have to analyze the lumbar spine shape as well as the mentioned lifestyle factors.
%% -------------------------------
\section{\uppercase{Related Work}}
\label{sec:RelatedWork}
\noindent \com{Our own work (VIS, VMV, BVM). Sylvias paper with reference to the methods. \textbf{More information necessary here!}}
\com{From VIS Paper}

ToDo Hier weitermachen
Steenwijk
Turkey (Skeweness-Paper)
Angelelli
Abschließend Uniqueness beschreiben

\paragraph{Visual Analysis of Heterogenous Non-Image Data.}
Zhang et al. \cite{Zhang2012} provide a web-based system for analyzing subject groups with linked views and batch-processing capabilities for categorizing new subject entries into the data set.
%
Their definition of a cohort differs from the understanding of the term in an epidemiological context by denoting every parameter-divided subject group as individual cohort.
%
Due to the short paper length, detail is missing on the data types and their algorithms of identifying similar subjects or whether they employ statistical measures.
%
We employ the idea of adding variables via drag and drop into a canvas area.

Generalized Pairs Plots (\texttt{GPLOM´S}) are an information visualization technique comparing heterogenous variables pairwise using a plot-matrix grouped by type \cite{GPLOMS, Francois2013}.
%
They are useful to gain an overview over numerous variables and their distributions.
%
Histograms, bar charts, scatter plots and heat maps are used to visualize variable combinations with regard to their type.
%
The resulting matrix provides an \emph{overview visualization}, but requires a lot of screen space for many variables (127 in our application scenario).
%
%Brushing, linking and filtering are feasible with \texttt{GPLOMS}, but have limitations such as making only one category brushable at a time.
%
%We applied this technique to our data and it shows promising potential for simultaneously visualizing many different variables. %but does not fit in the scope of this paper.
%
We incorporate the idea of adaptive type-dependent visualizations.
%\add{The inspiration on the chosen visualization techniques stems from this publication.}
%
Dai et al. \cite{Dai2005} explored risk factors by incorporating choropleth maps of epidemiological variables (e.g., mortality rates in a region) with parallel coordinates, bar charts and scatter plots with integrated regression lines.
%
Their findings yielded a \emph{Concept Map}, which linked cancer-related associations via graph edges.
%
While their goal to identify possible risk factors using socio-economic and health data is similar to ours, they focus on iteratively refining defined hypotheses and on geographical data.
%
We employ the use of small multiples for incorporating heterogenous data types for comparability.
%
Chui et al. \cite{Chui2011} visualized associations in time-dependent epidemiological data using time-series plots highlighting risk factor differences in age and gender.
%
%The connection of time series plots with histograms and heat maps allows for fast insight into complex relationships and 
While the work shows how different visualization techniques provide insight into these data sets, it focuses on the time aspect, which is not present for our data.

\paragraph{Prior Work.}
We visualized lumbar spine variabilities based on a semi-automatic shape detection algorithm of 490 participants of the \texttt{SHIP-2} cohort \cite{Klemm2013VMV}.
%
Hierarchical agglomerative clustering divided the population into shape-related groups.
%
As proof of concept, a relation between the size of the segmented shape and the measured size of the subjects was shown.
%
This work focuses on incorporating these derived data as new variables, enabling to include it into the hypothesis validation and generation process.
%
When applying clustering techniques to the non-image data it was found that \texttt{k-Prototypes} and \texttt{DBSCAN} are appropriate, but are strongly dependent on the chosen variables and distance measures \cite{Klemm2014BVM}.
%
Niemann et al. \cite{Niemann2014} presented an interactive data mining tool for the assessment of risk factors of hepatic steatosis, the fatty liver disease.
%
Association rules created by data mining methods can be analyzed interactively with their tool and highlight potentially overlooked variables.

@@

  \includegraphics[width=0.82\textwidth]{figures/results}
